Openness and computational reproducibility in plant pathology: where do we stand and a way forward

### 2022-10-13 analysis/paper/paper.Rmd
Address R2's comment Re: Order of statistical analysis
Clarify model structures

The statistical analysis was performed using `r version$version.string` [@RCT2022] on an Apple MacBook Pro (13-inch, M1, 2020).

@@

@@

Six models were applied, with half predicting the response of code availability and the other half predicting data availability.
Models 1 and 2 were used to predict responses with publishing journal, using the journals' abbreviation as the predictor the formula took the form: response \~ abbreviation (model 1: code_availability \~ abbreviation and model 2: data_availability \~ abbreviation) and the models included assignee as a random effect (formula: \~1 \| assignee).
Models 3 and 4 were used to predict the two responses with year as the predictor (formula: response \~ year) and the models included abbreviation and assignee as random effects (formula: list(\~1 \| abbreviation, \~1 \| assignee)).
Models 5 and 6 were used to predict the responses using IF_5year as the predictor the formula took the form: response \~ IF_5year and included assignee as random effect (formula: \~1 \| assignee) without journal as a random effect to mitigate correlation between impact factor and the journal of publication.

@@

@@

When predicting the availability of computational methods (code) and availability, all journals were compared with *Phytopathology*&reg; as the reference with latent scores having a mean of zero.

There were no clear differences that the analysis could detect between any of the journals as compared with *Phytopathology*&reg; for the sharing of computational methods (code); the effects of all parameters were undecided (Figure \@ref(fig:journal-parameters)A, Table S1, Figure S2).

When predicting the availability of data availability as compared with *Phytopathology*&reg;, publications in one journal, *Phytopathologia Mediterranea*, can be considered clearly more likely to share data than publications in *Phytopathology*&reg;, but no effects were detectable in any other journal title (\@ref(fig:journal-parameters)B, Table S2, Figure S3).

The effect of *Phytopathologia Mediterranea* (Median = 1.68, 95% CI [0.78, 2.57]) has a 99.99% probability of being positive (> 0), 99.98% of being significant (> 0.05), and 99.83% of being large (> 0.30). The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 19017).

@@

The methods for selecting articles meant that only page numbers between 1 and 150 were selected.
This approach assumed that there was no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year and some journals, _e.g._, *Phytopathology*&reg;&reg; have focus issues with invited authors.
In spite of some publications being from special issues, there were no detectable differences in the reproducibility of the articles as defined and measured by this article, so the assumption appears to be correct.

Except for a few isolated cases, most papers were not fully computationally reproducible.
Very few authors choose to share both data and code.
More authors shared data due to journal requirements to share sequence data, but other types of data related to field experiments or other laboratory studies were not likely to be shared.
Code sharing was extremely rare, but in cases where it was shared, most were included as a part of the journal article's publication as extra materials rather than through data repositories.
Authors publishing in *Phytopathologia Mediterranea* were more likely than articles published in *Phytopathology*&reg; to share data.
Several articles published in this journal either made the data available through GenBank or it was noted that all data were available through the article and associated supplementary materials.
The journal ranking highest for sharing data, *Phytopathologia Mediterranea*, did not have any articles that were sampled in 2018 (Table \@ref(tab:journals)).
There is no clear reason for the year 2018 to have more data shared than other years other than random chance that we found.
The reasons for not sharing code or data were not clearly available from the papers themselves and so this work was unable to determine possible reasons for this situation.
We recognize that in some cases there may be commercial or intellectual property (IP) reasons for not sharing data or code, but these reasons should be clearly stated.
However, in most cases, the data are collected with public funding and the code was developed using similar funding and there are no commercial or IP issues that preclude sharing and in fact the funding agency may have guidelines in place for sharing these materials.

>>>>

@@

Six models were applied, with half predicting the response of code availability and the other half predicting data availability.
Models 1 and 2 were used to predict responses with publishing journal, using the journals' abbreviation as the predictor (formula: response \~ abbreviation).
Models 3 and 4 were used to predict the two responses with year as the predictor (formula: response \~&nbsp;year) and the models included abbreviation and assignee as random effects (formula: list(\~1&nbsp;\|&nbsp;abbreviation,&nbsp;\~1&nbsp;\|&nbsp;assignee)).
Models 5 and 6 were used to predict the responses using IF_5year as the predictor, (formula: response&nbsp;\~&nbsp;IF_5year) and included assignee as random effect (formula:&nbsp;\~1&nbsp;\|&nbsp;assignee) without journal as a random effect to mitigate correlation between impact factor and the journal of publication.
All models included the assigned article evaluator as a random effect (formula:&nbsp;\~1&nbsp;\|&nbsp;assignee) to address inter-rater effects on scoring.

@@

@@

All statistical analysis was performed using `r version$version.string` [@RCT2022] on an Apple MacBook Pro (13-inch, M1, 2020).

@@

When predicting the availability of computational methods (code) and availability, all journals were compared with *Phytopathology*® as the reference with latent scores having a mean of zero.

There were no clear differences that the analysis could detect between any of the journals as compared with *Phytopathology*® for the sharing of computational methods (code); the effects of all parameters were undecided (Figure \@ref(fig:journal-parameters)A, Table S1, Figure S2).

When predicting the availability of data availability as compared with *Phytopathology*®, publications in one journal, *Phytopathologia Mediterranea*, can be considered clearly more likely to share data than publications in *Phytopathology*®, but no effects were detectable in any other journal title (\@ref(fig:journal-parameters)B, Table S2, Figure S3).

The effect of *Phytopathologia Mediterranea* (Median = 1.68, 95% CI [0.78, 2.57]) has a 99.99% probability of being positive (\> 0), 99.98% of being significant (\> 0.05), and 99.83% of being large (\> 0.30).
The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 19017).

@@

The methods for selecting articles meant that only page numbers between 1 and 150 were selected.
This approach assumed that there was no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year and some journals, *e.g.*, *Phytopathology*®® have focus issues with invited authors.
In spite of some publications being from special issues, there were no detectable differences in the reproducibility of the articles as defined and measured by this article, so the assumption appears to be correct.

Except for a few isolated cases, most papers were not fully computationally reproducible.
Very few authors choose to share both data and code.
More authors shared data due to journal requirements to share sequence data, but other types of data related to field experiments or other laboratory studies were not likely to be shared.
Code sharing was extremely rare, but in cases where it was shared, most were included as a part of the journal article's publication as extra materials rather than through data repositories.
Authors publishing in *Phytopathologia Mediterranea* were more likely than articles published in *Phytopathology*® to share data.
Several articles published in this journal either made the data available through GenBank or it was noted that all data were available through the article and associated supplementary materials.
The journal ranking highest for sharing data, *Phytopathologia Mediterranea*, did not have any articles that were sampled in 2018 (Table \@ref(tab:journals)).
There is no clear reason for the year 2018 to have more data shared than other years other than random chance that we found.
The reasons for not sharing code or data were not clearly available from the papers themselves and so this work was unable to determine possible reasons for this situation.
We recognize that in some cases there may be commercial or intellectual property (IP) reasons for not sharing data or code, but these reasons should be clearly stated.
However, in most cases, the data are collected with public funding and the code was developed using similar funding and there are no commercial or IP issues that preclude sharing and in fact the funding agency may have guidelines in place for sharing these materials.
